\vspace{-15mm} \fontsize{24pt}{10pt}\selectfont \textbf{Traversing Symmetric NAT with predictable port allocation function}

### 2013-12-11 paper.tex
evaluation, discussion, conslusion

% ------
% Math
\usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
\newtheorem{mydef}{Definition}
\newcommand{\ignore}[1]{}
% \setlength{\parskip}{0cm plus4mm minus3mm}

@@

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
respectively. By default an internal host does not known it's external address and NAT type thus technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} of the NAT traversal algorithm.

UDP hole punching works as follows:
\begin{compactitem}
 \item [1.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
Locally a~new mapping is created ($PORT^A_{nat}$), packet is dropped on B's~NAT.
 \item [2.] $B: \; IP^A_{nat}:PORT^A_{nat} \longleftarrow  IP^B_{in}:PORT^B_{in}$ \\
Locally a~new mapping is created ($PORT^B_{nat}$), packet reaches A using allocation created in step~1.
 \item [3.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
Packet reaches B using allocation created in in step~2.
\end{compactitem}

@@

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda) | \lambda > 0 \}$. For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively
\end{compactitem}
\end{mydef}
If third parameter is not important for the algorithm it can be ignored in notation.

@@

\begin{mydef}
Expected port value algorithm
\begin{align*}
& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
& \mathcal{A}(u, s, \lambda) = (0, s(1 + \lambda T))
\end{align*}
\end{mydef}

@@

\begin{mydef}
Poisson sampling algorithm 
\begin{align*}
& u \in \mathcal{U}, s \in \mathcal{S}, \lambda \in \mathcal{M}:\\
& \mathcal{A}(u, s, \lambda) = (0, \text{Poisson}(\lambda T (1+s C(\lambda)))
\end{align*} where $\text{Poisson}(\mu) = X \sim Po(\mu)$ samples Poisson distribution with
parameter $\mu$, function $C(\lambda)$ gives a constant 
coefficient\footnote{This function was found empirically by maximizing success rate}.
\end{mydef}

@@

\paragraph{Algorithm I and II comparison.} 
Algorithm I has following benefits over II:
\begin{compactitem}
 \item works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
 \item can be stopped as soon as connection is established, overhead is proportional to $\lambda$.
 \item single source port is used, only one listening thread is needed to implement
it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
This fact makes practical implementation simple and lightweight w.r.t. needed system resources.
 \item is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
\end{compactitem}

ToDo;
\begin{itemize}
 \item ToDo: algorithm success rate evaluation w.r.t. $\lambda, T$, graphs, graphs, graphs, \dots
 \item ToDo: algorithm pros/cons w.r.t. system requirements, applicability
 \item ToDo: fitting model to the real data from MU network
 \item ToDo: evaluating algorithms on real data from MU network
\end{itemize}

\section{Discussion}
ToDo; What can be done better, different settings, different $\lambda$ on both sides.

Success rate of the algorithm strongly relies on accurate measurements of the network 
properties the client resides in.  Usually

\section{Conclusion}
ToDo; New algorithm

>>>>

% ------
% Math
\usepackage{amsthm}
\usepackage{amsmath, mathtools}
\usepackage{fixltx2e}
% \usepackage{wrapfig}
\usepackage{tikz}
\usepackage{float}

\newtheorem{mydef}{Definition}
\newcommand{\ignore}[1]{}
% \setlength{\parskip}{0cm plus4mm minus3mm}

@@

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
respectively. By default an internal host does not known it's external address and NAT type thus technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\footnote{Since the algorithm itself does nothing} of the NAT traversal algorithm.

UDP hole punching works as follows:\\
\begin{compactitem}
 \item [1.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
Locally a~new mapping is created ($PORT^A_{nat}$), packet is dropped on B's~NAT.
 \item [2.] $B: \; IP^A_{nat}:PORT^A_{nat} \longleftarrow  IP^B_{in}:PORT^B_{in}$ \\
Locally a~new mapping is created ($PORT^B_{nat}$), packet reaches A using allocation created in step~1.
 \item [3.] $A: \; IP^A_{in}:PORT^A_{in} \longrightarrow IP^B_{nat}:PORT^B_{nat}$ \\
Packet reaches B using allocation created in in step~2.
\end{compactitem}

@@

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm
\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively
\end{compactitem}
\end{mydef}
If the third parameter is not important for the algorithm it can be ignored in the notation.

Prior algorithm run it is needed to 
determine NAT state (UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
starts from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
the NAT, e.g., difference between the last measurement and just before algorithm execution.

@@

\begin{mydef}
Expected port value algorithm
\begin{align*}
& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + s(1 + \lambda T))
\end{align*}
\end{mydef}

@@

\begin{mydef}
Poisson sampling algorithm 
\begin{align*}
& u \in \mathcal{U}, s \in \mathcal{S}, (\lambda, t_s) \in \mathcal{M}:\\
& \mathcal{A}(u, s, (\lambda, t_s)) = (0, \lambda t_s + \text{P}(\lambda T (1+s C(\lambda T)))
\end{align*} where $\text{P}(\mu) = X \sim Po(\mu)$, samples Poisson distribution with
parameter $\mu$, coefficient function $C(\lambda T)$ gives a constant 
coefficient\footnote{This function was found empirically by maximizing the success rate}.
Approximations of the $C(\lambda T)$ by an inverse logarithm and a quartic model:
\begin{align*}
C(x) = & ({0.163321 \cdot  ln(64.2568 \cdot x)})^{-1}\\
C(x) = & (0.172876+1.28162\cdot x-1.41256\cdot x^2 \\
       & +0.825093\cdot x^3-0.184726\cdot x^4)^{-1}
\end{align*}
\end{mydef}
% 
\subsection{Silent period in algorithms.} If the length of the silent period is too big
the algorithm can have problems with coping with a gap that occurred in NAT ports. Thus 
Algorithms III and IV estimates the starting offset caused by the silent period as an
expected value from $E[Po(\lambda \cdot t_1)] = \lambda\cdot t_1$.

@@

\paragraph{Algorithm I and II comparison.} 
Algorithm I has following benefits over II:
\begin{compactitem}
 \item Works with high success rate for $\lambda \in [0, 0.035] \sim 35$ new connections in 1~s on average.
 \item Can be stopped as soon as connection is established, overhead is proportional to~$\lambda$.
 \item Single source port is used, only one listening thread is needed to implement
it\footnote{NAT is abused to multiplex multiple connections to one source port} compared to the Alg. II.
Due to this fact a practical implementation is simple and lightweight w.r.t. needed system resources.
 \item Is $\lambda$-invariant. Provided network load is low no additional measurements are needed.
\end{compactitem}

\paragraph{Algorithms simulation.}

\begin{figure}[H]
% \begin{wrapfigure}{r}{0.5\textwidth}
% \begin{figure*}
% \begin{center}
% \leavevmode
% \centerline
%{\scalebox{0.40}{\includegraphics[width=\textwidth]{alg_1.pdf}}}
{\scalebox{0.43}{\includegraphics[trim=50 0 0 0]{alg_1.pdf}}}
% \end{center}
\caption{Algorithm evaluation in a simulation, $T=10~ms$. Number of a simulation rounds for one parameter setting is 1000.
 A, B, C correspond to the Algorithm I, II and III respectively.
 D, E correspond to the Algorithm IV with optimized $C(\lambda T)$ by an exhaustive search in D case, 
 in E case $C(\lambda T)$ is approximated by a polynomial.}
\label{fig:alg_eval}
\end{figure}
% \end{figure*}
% \end{wrapfigure}

\begin{figure}[H]
{\scalebox{0.43}{\includegraphics[trim=40 0 0 0]{alg_steps.pdf}}}
\caption{Steps of the algorithms needed to success on average w.r.t. $\lambda$. Algorithms denoted in a same way as in figure \ref{fig:alg_eval}}
\label{fig:alg_steps}
\end{figure}

In figure \ref{fig:alg_eval} is depicted result of a simulation of 
the algorithms, for different $\lambda$. Figure \ref{fig:alg_steps} depicts number of steps needed 
to succeed on average. From figures it is visible Alg. I is more suitable for networks with low 
workload since has $100\%$ success rate for $\lambda \in [0, 0.035]$. Outside this interval 
Alg. I is not working and another solution has to be applied. The problem is new connections are
created faster than algorithm progresses.

Alg. II is set to have fixed $\Delta=900$, in order to maximize the success rate. The disadvantage of this
algorithm is that it waits on a particular destination port and if it is taken by another connection
the algorithm fails.

Alg. IV has high success rate for 
noisy networks but is very sensitive on a $\lambda$ estimation what decreases practical usability 
if $\lambda$ changes quickly over time (non-homogeneous Poisson process). Closed form formula for $C(\lambda T)$ 
optimizing success rate and analytic derivation of the algorithm is still an open question.

\begin{itemize}
 \item ToDo: fitting model to the real data from MU network
 \item ToDo: evaluating algorithms on real data from MU network
\end{itemize}

\section{Discussion}
For simplicity here we assumed the $\lambda$ is the same on both sides, the new challenges 
can arise if we allow different parameters on both sides that we haven't studied yet.

Another open question is to find a model that optimizes given task, i.e., establishing a direct
connection. We proposed a start step, to model the same process that occurs on the network, i.e.,
Poisson process, but this method requires setting a coefficient function $C(\lambda T)$ for which 
we don't have exact analytic explanation. This method is also very sensitive to parameter fluctuations
and thus not very practical for a real world networks.

In our setting we assume $T=10~ms$ in order to minimize the speed of the evolution of a Poisson process.
Such short sending interval can cause some packets may get dropped on the path to the destination. But the 
important is whether they reach the local NAT in order to create a mapping in allocation table. As proposed 
in \citep{Wang:2006:RSN:1156422.1156550} it is possible to run the algorithm with $T=10~ms$ to create a mapping
and then re-run the algorithm with higher $T$, e.g., $T=100~ms$ in order to deliver packets to the destination
using created mapping and holes created in the previous round.

Success rate of the algorithm strongly relies on accurate measurements of the network 
properties the client resides in. They can be measured prior connection establishment 
from the point the client connects to the network, as is usually done in ICE and gather 
more information, build better network model and so on in this way.

Another problematic thing is if this algorithm is followed by a multiple host pairs on
the same network in the same time. They would artificially increase the network workload
causing it wouldn't follow Poisson distribution rendering our algorithms as not usable in
this case.

\section{Conclusion}
The aim of this work was to study problem of the NAT traversal from statistical point of view,
and to provide an algorithm with high success rate for networks with low workload. We proposed
several algorithms solving this problem but there are still some open questions remaining.
